%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CS624: Analysis of Algorithms
% Copyright 2015 Pejman Ghorbanzade <mail@ghorbanzade.com>
% Creative Commons Attribution-ShareAlike 4.0 International License
% More info: https://bitbucket.org/ghorbanzade/umb-cs624-2015s
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Question 7}
Prove that the following are equivalent.
\begin{enumerate}
  \item 
  \begin{enumerate}
    \item $V_1$ is a vertex cover of $G$.
    \item $V - V_1$ is an independent set in $G$.
  \end{enumerate}
  \item 
  \begin{enumerate}
    \item $V_2$ is an independent set in $G$.
    \item $V_2$ is a clique in $G^c$.
  \end{enumerate}
\end{enumerate}

\subsection*{Solution}

\begin{enumerate}
  \item To prove 1(a) and 1(b) are equivalent, we need to show vertex cover and independent set problems are polynomially equivalent.
  \begin{enumerate}
    \item Objective is to prove any independent set of vertices $V_2$ for an undirected graph $G = (V, E)$ is reducible to a vertex cover on $G' = (V_1, E')$ where $V_1 = V - V_2$.
    Since $V_1$ is an independent set, for any edge $e = (u,v) \in E$ either $u\in V_2$ or $v\in V_2$ which means either $u \in V_1$ or $v \in V_1$. Since $E$, by definition, contains all the edges in graph $G$, any $e \in E$ has at least one end in $V_1$, which means $V_1 = V - V_2$ is a vertex cover.
    \item Objective is to show any vertex cover $V_1 \in V$ in $G(V, E)$ is reducible to an independent set of vertices $V_2 = V - V_1$. Proof is given by contradiction that $V_2$ must form an independent set. Suppose $V_2$ is not an independent set which means there is an edge $e = (u,v) \in E$ such that both its endpoints are in $V_2$. Therefore $\{u, v\} \notin V_1$ which means set of vertices $V_1$ does not cover $e$ and therefore $V_1$ is not a vertex cover of $G$ which is a contradiction.
  \end{enumerate}
  Thus both problems 1(a) and 1(b) are polynomially reducible to each other which means they are equally hard.
  \item To prove 2(a) and 2(b) are equivalent, we need to show independent set and clique problems are polynomially equivalent.
  \begin{enumerate}
    \item Objective is to prove any independent set of vertices $V_2$ for an undirected graph $G = (V, E)$ is reducible to a clique on $G^c = (V, E')$. Since $V_2$ form an independent set, for any pair of vertices $\{u, v\} \in V_2$, $e = (u, v) \notin E$ which means $(u, v) \in E'$. Thus since all vertices in $V_2$ are mutually connected in $G^c = (V, E')$, they form a clique in $G'$.
    \item Objective is to prove any set of vertices $V_2$ that form a clique in $G^c=(V, E')$ form an independent set in $G = (V, E)$. Similar to previous proof, for any pair of vertices $\{u, v\}\in V_2$, $e = (u, v) \in E'$ which means $(u,v) \notin E$, i.e. no two vertices of $V_2$ are directly connected in $G=(V,E)$ which means they form an independent set.  
  \end{enumerate}
  Thus both problems 2(a) and 2(b) are polynomially reducible to each other which means they are equally hard.
\end{enumerate}
