%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CS637: Database-Backed Websites
% Copyright 2015 Pejman Ghorbanzade <pejman@ghorbanzade.com>
% Creative Commons Attribution-ShareAlike 4.0 International License
% More info: https://github.com/ghorbanzade/beacon
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Question 4}

\begin{enumerate}[label=(\alph*)]
\item Download \texttt{pizza1\_setup.zip} and unzip it, then copy it to \texttt{pizza1} for development.

\item Use \texttt{pizza1-setup}'s \texttt{database/createdb.sql} to load you own mysql \texttt{pizzadb}.
Also create a mysql user named \texttt{pizza\_user} for the app.
Report your commands and success or problems.
Show the resulting output of \texttt{showdb.sql}.

\item In \texttt{pizza1}, edit \texttt{database.php} to replace \texttt{xxxx} with your \texttt{cs.umb.edu} username.
(This will be used for executions on topcat.) Install pizza1 on your own Apache server at \texttt{htdocs/pizza1} and browse to \texttt{localhost/pizza1} to try it out.
Report on your success or problems.
It should be able to list the toppings, but not much else.

\item Set up a Netbeans projects for \texttt{pizza1-setup} and \texttt{pizza1}.
Then show the edit of database.php done in c.
as a file \texttt{diff}.by selecting both file versions in Projects, right-clicking and using \texttt{Tools>Diff}.
 Report on success or problems.
Also run the project from Netbeans and report on success or problems.
\end{enumerate}

\subsection*{Solution}

All instructions were carried out successfully.

\begin{enumerate}[label=(\alph*)]
\item \texttt{pizza1\_setup.zip} directory was successfully setup for development.

\item Resulting output of \texttt{showdb.sql} is a four-column row as shown in Table \ref{tab41}.

\begin{table}[H]\centering
\begin{tabular}{| c | c | c | c |}
\hline
next\_order\_id & next\_topping\_id & next\_size\_id & current\_day\\ 
\hline
1 & 1 & 1 & 1\\
\hline
\end{tabular}
\caption{Output of \texttt{showdb.sql}}\label{tab41}
\end{table}

\item \texttt{pizza1} is now accessible both from offline XAMPP server and from topcat server on CS department website.

\item Projects \texttt{pizza1-setup} and \texttt{pizza1} were successfully setup in Netbeans IDE.
\end{enumerate}
